Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-02-15 ms.tex
Explain why we are throwing away phase

The \ac{CRLB} says that, element by element,
%
\begin{equation}
    \cov{
        \left(
        \begin{array}{c}
            \hat{\rho} - \rho \\
            \hat{\gamma} - \gamma \\
            \hat{\tau} - \tau
        \end{array}
        \right)
    } \geq \mathcal{I}^{-1} = \left(
        \begin{array}{ccc}
            1 & 0 & 0 \\
            0 & \rho^2 \overline{\omega^2}/{\omega_\mathrm{rms}}^2 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 \\
            0 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 & \rho^2/{\omega_\mathrm{rms}}^2
        \end{array}
    \right)
\end{equation}
%
where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter and ${\omega_\mathrm{rms}}^2 = \overline{\omega^2} - \overline{\omega}^2$. Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{fairhurst:2009},
%
\begin{equation}\label{eq:timing-crlb}
    \std \left(\hat{\tau} - \tau \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho}{\omega_\mathrm{rms}}.
\end{equation}

@@

\marginpar{Talk here about likelihood structure---the `gaussian on a cylinder' concept, why we would want to project out the phase error.}%
%
A wrinkle is that $\mathcal{I}_{\tau\tau}$ depends on the true value of $\rho$. If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
%
\begin{eqnarray}
    \hat{\rho} - \rho \sim \mathcal{N}\, [0, 1] \\
    \hat{\gamma} - \gamma \sim \mathcal{N}\, \left[0, \overline{\omega^2} / {\omega_\mathrm{rms}}^2\right] \\
    \hat{\tau} - \tau \sim \mathcal{N} \left[0, 1 / {\omega_\mathrm{rms}}^2\right].
\end{eqnarray}

>>>>

The \ac{CRLB} says that, element by element,
%
\begin{equation}\label{eq:covariance-matrix}
    \cov{
        \left(
        \begin{array}{c}
            \hat{\rho} - \rho \\
            \hat{\gamma} - \gamma \\
            \hat{\tau} - \tau
        \end{array}
        \right)
    } \geq \mathcal{I}^{-1} = \left(
        \begin{array}{ccc}
            1 & 0 & 0 \\
            0 & \rho^2 \overline{\omega^2}/{\omega_\mathrm{rms}}^2 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 \\
            0 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 & \rho^2/{\omega_\mathrm{rms}}^2
        \end{array}
    \right)
\end{equation}
%
where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter and ${\omega_\mathrm{rms}}^2 = \overline{\omega^2} - \overline{\omega}^2$. Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{fairhurst:2009},
%
\begin{equation}\label{eq:timing-crlb}
    \std \left(\hat{\tau} - \tau \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho}{\omega_\mathrm{rms}}.
\end{equation}

@@

The likelihood in \Eref{eq:gaussian-likelihood} is the quantity that is maximized by the matched\nobreakdashes-filter estimator. However, just because we have the covariance matrix defined by \Eref{eq:covariance-matrix} does not mean that the distribution of estimator errors is multivariate Gaussian. Nonetheless, it is straightforward to qualitatively describe the distribution.

If we were to suppress the time parameter $\tau$, then we could reduce the likelihood to the form
%
\begin{equation}
    \mathcal{L}(z; \rho, \gamma) \propto \exp \left[ -\frac{1}{2} \left|z - e^{\imath \gamma} \rho \right|^2 \right]
\end{equation}
%
for a complex scalar measurement $z$. This likelihood also describes the distribution of the length $\rho$ and angle $\gamma$ of a vector whose Cartesian components are perturbed with normally distributed errors. For this problem, $\rho$ and $\gamma$ are uncorrelated, $\rho^2$ being noncentrally $\chi^2$\nobreakdashes-disributed with two degrees of freedom and $\gamma$ being von~Mises\nobreakdashes-distributed.
%
\marginpar{Picture here?}

\marginpar{What about if we suppress $\gamma$ or $\rho$? Do those cases teach us anything?}
%
We know that the matched\nobreakdashes-filter estimator's distribution, given a particular true value of $\rho$, $\gamma$, and $\tau$, must go to zero at $\hat{\rho}=0$ and must be periodic in $\gamma$ but not $\tau$. From just these properties, the estimator's distribution should be an approximate Gaussian, but wrapped diagonally around a cylinder where the radius represents $\hat{\rho}$, the angle around the cylinder represents $\hat{\gamma}$, and the distance up or down the cylinder represents $\hat{\tau}$.

There are many ways to realize distribution function with these properties, but they tend to be difficult to work with.
%
\marginpar{Give some examples.}
%
An alternative is to just discard the one periodic parameter, $\gamma$, entirely. We do not know the exact distribution of the remaining parameters $\rho$ and $\tau$, but we are less liable to make egregious errors by assuming that they are Gaussian than if we assumed that the estimation errors in $\gamma$ were Gaussian. Projecting out $\gamma$ gives us the covariance matrix
%
\begin{equation}
    \cov \left(
        \begin{array}{c}
            \hat{\rho} - \rho \\
            \hat{\tau} - \tau
        \end{array}
    \right) =
    \left(
        \begin{array}{cc}
            1 & 0 \\
            0 & \rho^2/{\omega_\mathrm{rms}}^2
        \end{array}
    \right).
\end{equation}

A wrinkle is that this covariance matrix still depends on the true, unknown value of $\rho$. If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can then write the distribution of the estimator error as
%
\begin{eqnarray}
    \hat{\rho} - \rho \sim \mathcal{N}\, [0, 1] \\
    \hat{\tau} - \tau \sim \mathcal{N} \left[0, \hat{\rho}^2 / {\omega_\mathrm{rms}}^2\right],
\end{eqnarray}
%
valid at moderate to high \ac{SNR}.
